---
title: "ex2"
output: github_document
date: '2022-05-16'
---
## Load objects
### Load Libraries
```{r }
library(tidyverse)
library(ggraph)
library(tidygraph)
library(igraph)
```

### Load nodes table
```{r }
nodes <- read_csv(file="exercices/Exercice 2_Nodes table.csv")
show(nodes)
head(nodes,10)
```
### Load links table
```{r }
links <- read_csv(file="exercices/Exercice 2_Links table.csv")
show(links)
head(links,17)
```
## Visualize network
```{r }
g <- graph_from_data_frame(d=links, vertices=nodes, directed=FALSE)
par(mar=c(0,0,0,0))

###  Change size of the nodes and node labels so that they match their importance
V(g)$size <- log(strength(g)) * 4 + 3

###  Change color of node based on seat choice
dark_side <- c("A", "B", "C", "D")
light_side <- c("1", "2", "3", "4", "5", "6")
V(g)$color[V(g)$name %in% dark_side] <- "red"
V(g)$color[V(g)$name %in% light_side] <- "gold"
vertex_attr(g)

plot(g, layout=layout_with_fr, main="Fakebook bus network graph")
show(g)

###  Add legend to indicate what the colors correspond to
legend(x=.75, y=.75, legend=c("Open seat", "Filled seat"), 
       pch=21, pt.bg=c("red", "gold"), pt.cex=2, bty="n")
```

## Calculate measures of centrality
```{r }
##  Degree centrality
degree <- degree(g,-4)
show(tail(degree,4))

## Nodes B, C, and D all have the same highest degree centrality (number of edges connected to it) and hence more influence than A

## Closeness centrality
closeness <- closeness(g)
show(tail(closeness,4))

## Nodes A and D all have the shortest closeness centrality (mean distance from a node to other nodes) and hence more direct influence on others

## Betweeness centrality

betweenness <- betweenness(g)
show(tail(betweenness,4))

## Node A has the highest betweenness centrality (extent to which a node lies on paths between other nodes) and hence more influence within a network, followed by nodes B and C
```

## Conclusion
### Choice of a seat: in the context of short bus rides, seat B is likely to be indicated because it would be beneficial for building a network. While for long bus rides, seat A seems to be the most desirable option because it would be beneficial for connecting people with others.