---
title: "Exercise 4"
output: github_document
date: '2022-05-30'
---

## Load and attach add-on packages
```{r }
library(readr)
library(arrow)
require(readr, lib.loc = NULL, quietly = FALSE,
        warn.conflicts,
        character.only = FALSE,
        mask.ok, exclude, include.only,
        attach.required = missing(include.only))
require(arrow, lib.loc = NULL, quietly = FALSE,
        warn.conflicts,
        character.only = FALSE,
        mask.ok, exclude, include.only,
        attach.required = missing(include.only))
```

#Load the files and add the following variables for examiners:
Load the following data: applications from `app_data_sample.parquet` + edges from `edges_sample.csv`
```{r }
applications <- read_parquet("/Users/justinenadeau-routhier/myrepo/myrepo/app_data_sample.parquet")
edges <- read_csv("/Users/justinenadeau-routhier/myrepo/myrepo/edges_sample.csv")
```
```{r }
applications
```
```{r }
edges
```
## Add the Gender variable for examiners:
```{r }
library(gender)
library(dplyr)

# List of examiner first names without repetitions
examiner_names <- applications %>%
  distinct(examiner_name_first)

examiner_names
```
## Get a table of examiner names and gender:
```{r }
library(purrr)
library (dplyr)
library(tidytext)
library(tidyr)

examiner_names_gender <- examiner_names %>%
  do(results = gender(.$examiner_name_first, method = "ssa")) %>%
  unnest(cols = c(results), keep_empty = TRUE) %>%
  select(
    examiner_name_first = name,
    gender,
    proportion_female)

examiner_names_gender
```
## Join  table back to  original applications data and discard the temporary tables we have just created to reduce clutter in our environment

``` {r }
# remove extra colums from the gender table
examiner_names_gender <- examiner_names_gender %>% 
  select(examiner_name_first, gender)

# joining gender back to the dataset
applications <- applications %>% 
  left_join(examiner_names_gender, by = "examiner_name_first")

# cleaning up
rm(examiner_names)
rm(examiner_names_gender)
gc()
```

## Join the data back to the applications table

## Add the Race variable for examiners:
```{r }
library(wru)

# List of examiner surnames without repetitions
examiner_surnames <- applications %>% 
  select(surname = examiner_name_last) %>% 
  distinct()

examiner_surnames
```
``` {r }

# Estimate likely race of an examiner and get probability across five broad US Census categories: white, black, Hispanic, Asian and other
examiner_race <- predict_race(voter.file = examiner_surnames, surname.only = T) %>% 
  as_tibble()

examiner_race
```

``` {r }

# Pick the race category that has the highest probability for each last name and then join the table back to the main applications table
examiner_race <- examiner_race %>% 
  mutate(max_race_p = pmax(pred.asi, pred.bla, pred.his, pred.oth, pred.whi)) %>% 
  mutate(race = case_when(
    max_race_p == pred.asi ~ "Asian",
    max_race_p == pred.bla ~ "black",
    max_race_p == pred.his ~ "Hispanic",
    max_race_p == pred.oth ~ "other",
    max_race_p == pred.whi ~ "white",
    TRUE ~ NA_character_))

examiner_race
```
## Join the data back to the applications table
``` {r }
# removing extra columns
examiner_race <- examiner_race %>% 
  select(surname,race)

applications <- applications %>% 
  left_join(examiner_race, by = c("examiner_name_last" = "surname"))

rm(examiner_race)
rm(examiner_surnames)
gc()
```

## Add the Tenure variable for examiners:
```{r }
library(lubridate) # to work with dates

examiner_dates <- applications %>% 
  select(examiner_id, filing_date, appl_status_date) 

examiner_dates
```

``` {r }
examiner_dates <- examiner_dates %>% 
  mutate(start_date = ymd(filing_date), end_date = as_date(dmy_hms(appl_status_date)))
```

``` {r }
examiner_dates <- examiner_dates %>% 
  group_by(examiner_id) %>% 
  summarise(
    earliest_date = min(start_date, na.rm = TRUE), 
    latest_date = max(end_date, na.rm = TRUE),
    tenure_days = interval(earliest_date, latest_date) %/% days(1)
    ) %>% 
  filter(year(latest_date)<2018)

examiner_dates
```
## Join the data back to the applications table
``` {r }
applications <- applications %>% 
  left_join(examiner_dates, by = "examiner_id")

rm(examiner_dates)
gc()
```

``` {r }
head(applications)
```

# Pick two workgroups to focus on (represented by the first 3 digits of `examiner_art_unit` value): 175 and 176 
``` {r }
colummn_7 <- applications %>% select(examiner_art_unit)

#Select the first two workgroups: 175 and 176
head(unique(colummn_7),2) 
```

## List examiners from the two selected workgroups (175 and 176)
``` {r }

# String staring from the 1st number to 3rd number is extracted as a substring
workgroup_175 = applications[substr(applications$examiner_art_unit, 1,3)==175,]
workgroup_176 = applications[substr(applications$examiner_art_unit, 1,3)==176,]

head(workgroup_175)
head(workgroup_176)
```
## Show how  the two workgroups (175 and 176) compare on examiners’ demographics (summary statistics about gender)
``` {r }
# Calculate gender ratio in the workgroups 175
round(table(workgroup_175['gender'])/dim(workgroup_175)[1]*100,2)
```
From the output we can observe:
31.13% of examiners in the workgroup 175 are female and
58.08% of examiners in the workgroup 175 are male. Therefore, the workgroup 175 is significantly male-dominated.
``` {r }
# Calculate gender ratio in the workgroups 176
round(table(workgroup_176['gender'])/dim(workgroup_176)[1]*100,2)
```
From the output we can observe:
30.72% of examiners in the workgroup 176 are female and
58.62% of examiners in the workgroup 176 are males. Therefore, the workgroup 176 is significantly male-dominated.
``` {r }
# Calculate average gender ratio in the applications table (as a reference)
round(table(applications['gender'])/dim(applications)[1]*100,2)
```
From the output we can observe:
28.30% of examiners in the applications table  are female and
56.65% of examiners in the applications table are males. Therefore, the average examiners is significantly male-dominated.

## Show how the two workgroups (175 and 176) compare on examiners’ demographics (summary plots about gender)
``` {r }
library(gridExtra)
library(tidyverse)

plot_gender_175 <- ggplot(data=workgroup_175, aes(x=gender)) +
  geom_bar(aes(y = (..count..)/sum(..count..)*100),color="black",fill="light green") + 
  theme_minimal() +
  ylab("Ratio (%)")+
  xlab("Gender")+
  ylim(0,70)+
  ggtitle(paste0("Gender Ratio for Workgroup 175"))

plot_gender_176 <- ggplot(data=workgroup_176, aes(x=gender)) +
  geom_bar(aes(y = (..count..)/sum(..count..)*100),color="black",fill="light blue") +
  theme_minimal() +
  ylab("Ratio (%)")+
  xlab("Gender")+
  ylim(0,70)+
  ggtitle(paste0("Gender Ratio for Workgroup 176"))

plot_gender_avg <- ggplot(data=applications['gender'], aes(x=gender)) +
  geom_bar(aes(y = (..count..)/sum(..count..)*100),color="black",fill="light grey") +
  theme_minimal() +
  ylab("Ratio (%)")+
  xlab("Gender")+
  ylim(0,70)+
  ggtitle(paste0("Average Gender Ratio for Examiners"))

grid.arrange(plot_gender_175,plot_gender_176,plot_gender_avg,widths=c(1,1))
```

## Show how the two workgroups (175 and 176) compare on examiners’ demographics (summary statistics about race)
``` {r }
# Determine racial profile in the workgroups 175
race_175 <- round(table(workgroup_175['race'])/dim(workgroup_175)[1]*100,2)
race_175
```
From the output we can observe:
19.27% of examiners in the workgroup 175 are Asian,
1.79% of examiners in the workgroup 175 are black,
1.71% of examiners in the workgroup 175 are Hispanic, and
77.23% of examiners in the workgroup 175 are white. Therefore, the workgroup 175 is significantly white-dominated.
``` {r }
# Determine racial profile in the workgroups 176
race_176 <- round(table(workgroup_176['race'])/dim(workgroup_176)[1]*100,2)
race_176
```
From the output we can observe:
25.11% of examiners in the workgroup 176 are Asian,
3.35% of examiners in the workgroup 176 are black,
3.01% of examiners in the workgroup 176 are Hispanic, and
68.53% of examiners in the workgroup 176 are white. Therefore, the workgroup 176 is significantly white-dominated.
``` {r }
# Determine racial profile in the applications table (as a reference)
race_avg <- round(table(applications['race'])/dim(applications)[1]*100,2)
race_avg
```
From the output we can observe:
29.18% of examiners in the workgroup 175 are Asian,
4.46% of examiners in the workgroup 175 are black,
2.96% of examiners in the workgroup 175 are Hispanic,
63.38% of examiners in the workgroup 175 are white, and
0.02% of examiners in the workgroup 175 are a member of another race.Therefore, the average examiners is significantly white-dominated.

## Show how the two workgroups (175 and 176) compare on examiners’ demographics (summary plots about race)
``` {r }
library(gridExtra)
library(tidyverse)

par(mfrow=c(1,3)) 
lbls <- c("Asian", "black", "Hispanic", "white")
lbls_o <- c("Asian", "black", "Hispanic", "other", "white")

plot_race_175 <- pie(race_175,labels = lbls, col=rainbow(length(race_175)),main = "Racial Profile in Workgroups 175",clockwise = TRUE)

plot_race_176 <- pie(race_176,labels = lbls, col=rainbow(length(race_175)),main = "Racial Profile in Workgroups 176",clockwise = TRUE)

plot_race_avg <- pie(race_avg,labels = lbls_o, col=rainbow(length(race_avg)),main = "Average Racial Profile Examiners",clockwise = TRUE)
```

# Create advice advice_networks from `edges_sample`

``` {r }
library(dplyr)
library(tidygraph) 

# Identify the workgroups of interest
art_unit = distinct(subset(applications, select=c(examiner_art_unit, examiner_id)))

# Add workgroups to use  as an indicator in the graph and select the groups of interest
art_unit$work_group = substr(art_unit$examiner_art_unit, 1,3)
art_unit = art_unit[art_unit$work_group==175 | art_unit$work_group==176,]

# Merge
merger = merge(x=edges, y=art_unit, by.x="ego_examiner_id", by.y="examiner_id", all.x=TRUE) %>% rename(art_unit_ego = examiner_art_unit, work_group_ego=work_group)

# Drop the NA values (other groups than 175 or 176)
merger = drop_na(merger)

# Repete for the alters
merger = merge(x=merger, y=art_unit, by.x="alter_examiner_id", by.y="examiner_id", all.x=TRUE)%>% rename(art_unit_alter = examiner_art_unit, work_group_alter=work_group)

merger = drop_na(merger)

```

There are 1651 edges left  representing examiners from the workgroups 175 and 176.

``` {r }
# Define focal nodes ("egos") and the nodes to whom egos are directly connected to ("alters") to build the Ego advice_networks
library(igraph)
egos = subset(merger, select=c(ego_examiner_id, art_unit_ego, work_group_ego)) 
egos = egos %>% rename(examiner_id=ego_examiner_id,art_unit=art_unit_ego,work_group=work_group_ego)

nodes = subset(merger, select=c(alter_examiner_id, art_unit_alter, work_group_alter)) 
nodes = nodes %>% rename(examiner_id=alter_examiner_id,art_unit=art_unit_alter, work_group=work_group_alter)

nodes = distinct(rbind(egos, nodes)) %>% group_by(examiner_id) %>% summarise(examiner_id = first(examiner_id), art_unit = first(art_unit), work_group = first(work_group))

# Creates an igraph graph from two data frames containing the edge list and edge/vertex attributes
advice_network <- graph_from_data_frame(merger, directed=TRUE, vertices=nodes)
print(advice_network, e=TRUE, v=TRUE)
```

``` {r }
library(ggraph)
# Calculate centrality scores in selected workgroups
# degree centrality (number of edges connected to it)
degree_examiners <- round(degree(advice_network, v=V(advice_network)),2)
# closeness centrality (mean distance from a node to other nodes)
closeness_examiners <- round(closeness(advice_network),2)
# betweenness centrality (extent to which a node lies on paths between other nodes) 
betweenness_examiners <- round(betweenness(advice_network),2) 

V(advice_network)$size = degree_examiners
V(advice_network)$bet = betweenness_examiners
V(advice_network)$clo = closeness_examiners
V(advice_network)$color = nodes$art_unit

ggraph(advice_network, layout="kk") +
  geom_edge_link()+
  geom_node_point(aes(size = degree_examiners, colour = closeness_examiners))
  coord_fixed()
```
# Calculate centrality scores for examiners in the selected workgroups
``` {r }
# Calculate three indices (degree, betweenness and closeness) to the advice_network and 
calculate_centralities = data.frame(cbind(degree_examiners, betweenness_examiners, closeness_examiners))

# Display the top 6 examiners based on calculated centrality indices
head(calculate_centralities[order(-degree_examiners),])
```
It can be observed that the `examiner 67829` has the highest degree of centrality and the highest associated closeness. The `examiner 60203` should also be considered due to it’s high betweenness, combined with a high degree of centrality. The relationship between centrality and other examiners’ characteristics has yet to be determined graphically.

# Create variable for application processing time ‘app_proc_time’ that measures the number of days (or weeks) from application filing date, until the final decision on it (patented or abandoned)
## Final decision date is computed as either abandon date or patent issue date
``` {r }
library(dplyr)
library(tidyselect)
application_time <- applications %>% 
    mutate(decision_date = coalesce(abandon_date,patent_issue_date)) %>%
    select(application_number,filing_date, abandon_date, patent_issue_date, decision_date, examiner_id, examiner_art_unit, gender, race, tenure_days) %>%
    filter(!is.na(decision_date))

head(application_time)
```
## Calculate the application processing time as the difference between filing date and final decision date
``` {r }
application_time <- application_time %>% 
    mutate(app_proc_time = difftime(decision_date, filing_date, units = "days"))

head(application_time)
```

# Estimate the relationship between centrality and `app_proc_time`

# Calculate average application processing time
``` {r }
# columns <- c("examiner_id", "gender", "race", "tenure_days", "app_proc_time")
# 
# examiners <- application_time %>%
#     group_by(across(all_of(columns))) %>%
#     dplyr::summarize(avg_app_proc_time = mean(app_proc_time, na.rm=TRUE), n_app = n()) %>%
#     drop_na()
```

# Run linear regression to determine relationship between centrality and app_proc_time
``` {r }
# lin_reg = lm(application_time,c('app_proc_time') ~ degree_examiners+ betweenness_examiners+ closeness_examiners + race + gender + tenure_days , data=applications)
# summary(lin_reg)
```